---
title: "Integrate Andrew's data"
output: html_document
---

The purpose of this script is to:  

* integrate the data that Andrew cleaned into the Cloud SQL database  
* identify the extent of the data, after which we need to automate the cleaning process  


First, let's load the relevant libraries.
```{r}
library(RSQLite)
library(DBI)
library(tidyverse)
library(stringr)
library(readr)
library(lubridate)
```

Upload the clean historical data, and remove erroneous super large values.
```{r}
# read data
cd <- read_csv("C:/Users/rpauloo/Documents/GitHub/cosumnes_shiny/clean/andrew_clean_data/LevelSenderAllData.csv")

# gather 
cd <- gather(cd, well, level, -Date) %>%  
  rename(date = Date) %>%                 # rename date column
  mutate(date = mdy_hm(date),             # convert date to datetime
         level = as.numeric(level)) %>%   # convert level to numeric
  filter(level < 1000)                    # filter erroneous large values
```

Find extent of the clean data. Looks like it's October 2, 2018.
```{r}
# visualize
cd %>% ggplot(aes(date, level, color = well)) + geom_line()

# find extent
cd %>% arrange(date) %>% tail(100) #%>% distinct(well)

# spread back to wide format
cd <- spread(cd, well, level)
```


Connect to our .sqlite database via R, and write the clean data to it.
```{r}
# password for gw obs db
pw <- read_rds("C:/Users/rpauloo/Documents/GitHub/cosumnes_shiny/dashboard/data/pw.rds")

# connect to the cloud SQL database 
db <- dbConnect(RMySQL::MySQL(),
                user = "gw_observatory",
                password = pw,
                host = "169.237.35.237",
                dbname = "gw_observatory",
                port = 33306)

# write clean data to database
dbWriteTable(db, name = "clean_historical_data_through_october", value = cd)

# list tables to ensure data has been written
dbListTables(db)
```


Wite elevation data for sites to cloud database.  
```{r}
# load elevation data for sites
elev <- read_csv("C:/Users/rpauloo/Documents/GitHub/cosumnes_shiny/clean/andrew_clean_data/elev.csv")

# password for gw obs db
pw <- read_rds("C:/Users/rpauloo/Documents/GitHub/cosumnes_shiny/dashboard/data/pw.rds")

# connect to the cloud SQL database 
cdb <- dbConnect(RMySQL::MySQL(),
                 user = "gw_observatory",
                 password = pw,
                 host = "169.237.35.237",
                 dbname = "gw_observatory",
                 port = 33306)

# write elevation data to the cloud database
dbWriteTable(cdb, name = "elevation_data", value = elev)
```


Disconnect from SQLite and cloud SQL databases.
```{r}
# disconnect
dbDisconnect(db)
dbDisconnect(cdb)
```

