---
title: "Using RSQLite to query LevelSender"
output: html_notebook
author: Rich Pauloo & Amy Yoder
date: February 21, 2018
---

The purpose of this script is to obtain data from the Home Station Database for Level Sender, in `C:/Users/rpauloo/Documents/LevelSender/db/levelsender.sqlite`. Ultimately, the goal is to develop an R Shiny Web application that automatically downloads the most up-to-date levelsender data and displays it.  

Information from Solinist on LevelSender can be found [here](https://www.solinst.com/products/dataloggers-and-telemetry/solinst-telemetry-systems/9500-levelsender/instructions/user-guide/5-3-0-data-from-software.php#5-4-0).  

The R interface to SQLite is documented [here](https://cran.r-project.org/web/packages/RSQLite/RSQLite.pdf), and offers far superior examples compared to anything else available on the web at the time of this writing.  


First, let's load the relevant libraries.
```{r}
library(RSQLite)
library(DBI)
library(tidyverse)
library(stringr)
```

Refresh emails and append to sqlite database.
```{r}
# simulates clicking the "Retrieve Emails Button"
system(
  shQuote(
    "C:/Program Files (x86)/Solinst/LevelSender/LSEmailClient.exe", 
    type = "cmd" # change to "sh" for Unix/bash, and "csh" for C-shell
  )
)

# opens the software which finalizes the data append from the emails that were read during the LSEmailClient.exe
# system(
#   shQuote(
#     "C:/Program Files (x86)/Solinst/LevelSender/LSender.exe",
#     type = "cmd" # change to "sh" for Unix/bash, and "csh" for C-shell
#   )
# )
```


Second, let's connect to our .sqlite database via R, and list the tables of data within.
```{r}
# connect to SQLite database that Solinist regularly updates
db <- dbConnect(SQLite(), dbname = "C:/Users/rpauloo/Documents/LevelSender/db/levelsender.sqlite")

# list of tables
db_tables <- dbListTables(db)
```

For now, we're interested in all of this data, so let's read it all in a list, and explore.
```{r}
# We can read in one table by simply naming it...
dbReadTable(db, "ConfigEmailSetup")

# ...or we can all the tables into a list...
table_list <- lapply(db_tables, function(x) { dbReadTable(db, x) } )

# ...and example them one-by-one.
table_list[[8]] 

# It looks like the "ReceivedEmail" table contains the groundwater level
dbReadTable(db, "ReceivedEmail")

# There's a lot of information that we don't need. Let's select what we do need.
dbReadTable(db, "ReceivedEmail") %>% 
  select(ReceivedDate, Subject, Body)
```

Looks like there are a lot of emails that aren't actually reports. Let's filter out the emails that are relevant to us, and give the columns more intuitive names.
```{r}
# read, select interesting data, filter for relevant emails with data, and rename columns
d <- dbReadTable(db, "ReceivedEmail") %>% 
  select(ReceivedDate, Subject, Body) %>% 
  filter( grepl("LS Report", Subject) ) %>% 
  rename(date = ReceivedDate, subject = Subject, body = Body)

# We want to arrange these emails by the date they were received, but first we need to convert the `Date` from a character vector to a `Date` object.
d$date <- as.POSIXct( strptime( d$date, "%Y-%m-%d %H:%M:%S" ) )

# Now we can arrange by date.
#d %>% arrange(date)
```

Oh no! Although these data are arranged by time, we're actaully looking at data from multiple wells! In the body of each observation is a serial number. Let's extract that so we can arrange by it as well before we mine the body of each email for data.


The body of the emails is complete spaggetti. We need to organize these and write rules to extract the data we need.
```{r}
# clean up one email 
strsplit(d$body[143], "\r\n")
```

Create a new 4 column dataframe (datetime, temp, level, id) for each email body, and store in a list. Keep in mind that some email bodies will have more than one level logger. We do this for one email body to begin with.
```{r}
l <- unlist(strsplit(d$body[100], "\r\n")) # email body as character vector

# get MW serial # 
id <- l[str_detect(l, "Serial: ")][2] # 1st serial # == level sender
id <- as.numeric(substr(id, 9, nchar(id)))

# get all datetime, temp, level data
i <- str_detect(l, "^[:digit:]") # all lines that begin with a digit
ld <- l[i] 
ld

# organize into a dataframe
m <- str_split_fixed(ld, ", ", 3) # matrix of strings
m[, 2:3] <- round(as.numeric(m[, 2:3]), 2)
df <- as.data.frame(m)
colnames(df) <- c("dt", "temp", "level")
df$dt <- as.POSIXct( strptime(df$dt, format = '%d/%m/%Y %H:%M:%S') ) # format dates
df$id <-  id # add id
df
```

Now we generalize the above code for all monitoring wells... but first check to see which records contain MW5.
```{r}
mw5 <- ": 283687" # serial number of MW5 (which includes barrologger)

# which records include MW5?
temp <- sapply(d$body, str_detect, mw5)
names(temp) <- NULL
#temp
sum(temp) # 182 records contain MW5

# number of serial numbers per email
temp2 <- lapply(d$body, function(x){unlist(strsplit(x, "\r\n"))} )
temp3 <- sapply(temp2, function(x){length(x[str_detect(x, "Serial: ")])})
data.frame(mw_5 = temp, n_serial = temp3) # visualize 
temp4 <- temp2[temp3==1] # Oneto Denir well doesn't have a serial number...
length(temp4) # sizteen records from Oneto Denir
```

Generalize for all monitoring wells.
```{r}
# strat with temp2, the formatted lines of each email body
lines <- lapply(d$body, function(x){unlist(strsplit(x, "\r\n"))} )

# serial number for mw5
mw5 <- ": 283687"

# function to apply
get_data <- function(v){ 
  
  #v  = lines[[1]]
  #v2 = lines[[2000]]
  baro <- NULL # initalize baro vector
  ss <- sum(str_detect(v, mw5)) # does mw5 appear in the email?
  
  # if well is MW 5
  if (ss == 1) {
    id <- v[str_detect(v, "Serial: ")][2]   # 2nd serial is level logger
    baro <- v[str_detect(v, "Serial: ")][3] # 3rd serial is baro  logger
  }

  # if the well is not mw5
  if (ss == 0) {
    id <- v[str_detect(v, "Serial: ")][2] # 2nd serial number is level logger
  }

  # subset for the levellogger serial number
  id <- as.numeric(substr(id, 9, nchar(id)))
  
  # if barometric pressure logger is present, get its serial
  if(!is.null(baro)){baro <- as.numeric(substr(baro, 9, nchar(baro)))}
  
  
  # get all datetime, temp, level data
  i <- str_detect(v, "^[:digit:]") # all lines that begin with a digit
  ld <- v[i] # subset for these lines
  
  # organize into a dataframe
  m <- str_split_fixed(ld, ", ", 3)          # matrix of strings
  m[, 2:3] <- round(as.numeric(m[, 2:3]), 2) # round temp and level
  df <- as.data.frame(m)                     # convert to df
  colnames(df) <- c("dt", "temp", "level")   # rename columns
  df$dt <- as.POSIXct(strptime(df$dt, format = '%d/%m/%Y %H:%M:%S')) # format dates
  df$id <-  id # add id
  df
  
  
}

get_data(d$body)
```

Convert from PSI to meters
```{r}
# convert from PSI to meters
psi_to_m <- function(psi){
  return(psi * 0.703070)
}
```





